String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-04-26 2023-1-19/sections/algorithms.tex
algorithm

As mentioned in Section \ref{sec:overview}, the main idea of our approach is to model the counting operators symbolically by registers in automata, instead of unfolding them explicitly. Additionally, we represent string lengths using registers. This reduces the satisfiability of string constraints that involve regex-counting and string-length to the satisfiability of linear integer arithmetic, which can be solved by off-the-shelf SMT solvers. We also propose techniques to reduce automata sizes and utilize under approximations to enhance performance.
\subsection{Encode Counting Operator to CEFA} \label{subsec:regex2cefa}
One key point of our approach is the encoding of counting operator. In this section, we define the construction of CEFA from the regular expression with counting. We divide the counting operator into \emph{handled} and \emph{non-handled}. The counting operator is called \emph{non-handled} if it is the sub-regex of other complement operator (e.g., $(\regex^{\{m,n\}})^C$), closure operator(e.g., $(\regex^{\{m,n\}})^*$), or counting operator (e.g., $(\regex^{\{m,n\}})^{\{m', n'\}}$), otherwise it is called \emph{handled}. A regex is called \emph{non-handled} if it contains any non-handled counting operators, and is called \emph{handled} otherwise.
% We must syntactically rewrite bounded repetition $\regex\{m,n\}$ if it is the sub-regex of complement (e.g., $(\regex\{m,n\})^C$), closure(e.g., $(\regex\{m,n\})^*$), and bounded repetition(e.g., $(\regex\{m,n\})\{m', n'\}$), we unwind it to $\regex^m\mid\cdots\mid\regex^n$. After this syntactic rewriting, we call the resulting regex $\regex'$ \emph{non-nested}.
Given a handled regex $\regex$, operations such as intersection, union, concatenation, complement, closure, and counting may appear. To construct the CEFA of $\regex$, we first create CEFA for each sub-regex and then combine them. Similar to the construction of NFA from regex\cite{aut_hopcraft}, our construction is inductive on the size of $\regex$. The base case is the construction for a single character, an empty string, or an empty language. The inductive step is the construction for the concatenation, union, intersection of two automata of sub-regexes, counting, closure, and complement of one automaton of sub-regex. We only discuss the inductive step for counting because it is the most important and the other operations are trivial. \newline
For a handled regular expression $\regex\{m,n\}$, suppose the CEFA recognizing $\regex$ is $\aut = (Q,\Sigma,\delta,q_I,F,\emptyset,\top)$, then the CEFA recognizing $\regex\{m,n\}$ is defined as $\aut_{m,n} = (Q, \Sigma, \delta',q_ I, F, r, \theta)$ where:
\begin{itemize}
  \item $r$ is a new register,
  \item $\delta'$ is composed by transition $q\xrightarrow[(0)]{a} q'$ for each transition $q\xrightarrow[()]{a} q' \in \delta$, and transition $q_f\xrightarrow[(1)]{\epsilon} q_I$ for each accepting state $q_f\in F$,
  \item $\theta$ is the function mapping each accepting state to a same linear integer arithmetic $m\leq r\leq n$.
\end{itemize}
The construction is illustrated in Fig.\ref{fig:construct_repetition}, where a new register $r$ is added to store the counting times. The transition $q\xrightarrow[(1)]{\epsilon} q_I$ is used to update the counting times for each $q\in F$. The accepting condition $\theta'$ constrains the counting times of the accepting word to the range $[m, n]$. 
\begin{figure}[h]
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \begin{tikzpicture}[
      shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
      initial text=, every state/.style={minimum size = 0.001cm},
      accepting text=$\top$, accepting/.style=accepting by arrow,
      ]

      \node[state,initial]            (q_0)                      {};
      \node[state, accepting]         (q_1) [right=of q_0]       {};
      \node [fit=(q_0) (q_1)] {$\cdots$};
    \end{tikzpicture}
    \caption{The CEFA recognizing $\regex$}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \begin{tikzpicture}[
      shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
      initial text=, every state/.style={minimum size = 0.001cm},
      accepting text=\myemph{$m\leq r\leq n$}, accepting/.style=accepting by arrow,
      ]

      \node[state,initial]            (q_0)                      {};
      \node[state, accepting]         (q_1) [right=of q_0]       {};
      \path[->] (q_1) edge[bend right]   node    [above] {\myemph{$\mathbf{\epsilon/(1)}$}} (q_0);
      \node [fit=(q_0) (q_1)] {$\cdots$};
    \end{tikzpicture}
    \caption{The CEFA recognizing $\regex^{\{m,n\}}$}
  \end{subfigure}
  \caption{The construction of counting operator}
  \label{fig:construct_repetition}
\end{figure}
\begin{example}
  To construct the CEFA recognizing $(ab)\{1,100\}$. As shown in Fig.\ref{subfig:rep_aut_ab1_100}, the base cases are the constructions of character $a$ (Fig.\ref{subfig:rep_aut_a}) and $b$ (Fig.\ref{subfig:rep_aut_b}). The first inductive step is to concatenate them to one automaton (Fig.\ref{sub@subfig:rep_aut_ab}). The second inductive step is to add counting information on the automaton. To do that we add a new register $r_1$ and a transition $q_2\xrightarrow[(1)]{\epsilon} q_0$ to update the repetition times. The accepting condition in accepting state is $1\leq r_1\leq 100$ so that the counting time of the accepting word is in the range $[1,100]$. After that we get the CEFA recognizing $(ab)\{1,100\}$ shown in Fig.\ref{subfig:rep_aut_ab1_100}.
  \begin{figure}[h]
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=$\top$, accepting/.style=accepting by arrow,
        accepting where=above
        ]

@@

        \path[->] (q_0) edge              node      [above]           {$a$/()} (q_1);
      \end{tikzpicture}
      \caption{The CEFA recognizing $a$}
      \label{subfig:rep_aut_a}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=$\top$, accepting/.style=accepting by arrow,
        accepting where=above
        ]

@@

        \path[->] (q_0) edge              node      [above]           {$b$/()} (q_1);
      \end{tikzpicture}
      \caption{The CEFA recognizing $b$}
      \label{subfig:rep_aut_b}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=$\top$, accepting/.style=accepting by arrow,
        accepting where=above
        ]

@@

        \path[->] (q_0) edge              node      [above]           {$a$/()} (q_1)
        (q_1) edge              node      [above]           {$b$/()} (q_2);
      \end{tikzpicture}
      \caption{The CEFA recognizing $ab$}
      \label{subfig:rep_aut_ab}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=\myemph{1}$\leq r_1\leq$\myemph{100}, accepting/.style=accepting by arrow,
        accepting where=above
        ]

@@

        \path[->] (q_0) edge              node      [above]           {$a$/(\myemph{0})} (q_1)
        (q_1) edge              node      [above]           {$b$/(\myemph{0})} (q_2)
        (q_2) edge [bend left]  node      [below]           {$\epsilon$/(\myemph{1})} (q_0);
      \end{tikzpicture}
      \caption{The CEFA recognizing $(ab)\{\myemph{1},\myemph{100}\}$}
      \label{subfig:rep_aut_ab1_100}
    \end{subfigure}
    \caption{The example of constructing regex with counting}
  \end{figure}
\end{example}
Above discusses the construction of counting operator in a handled regex. The counting operator of unhandled regex can not be constructed by the same way. For example, for an unhandled regex $(ab^{\{1,100\}})*$, there are infinite repetitions of regex $ab^{\{1,100\}}$. The counting times of each repetition are irrelevant, so that we can not represent them use finite registers. The way we solved this problem is to syntactically rewrite the unhandled counting operator. More exactly, the sub-regex $\regex^{\{m,n\}}$ in the unhandled regex is rewritten into $\underbrace{\regex\cdot\regex\cdots\regex}_{m \text{ times}}\mid\cdots\mid \underbrace{\regex\cdot\regex\cdots\regex}_{n \text{ times}}$.
% Based on the decision procedure defined in the paper \cite{atva2020}, we propose an efficient algorithm to solve ESL formula $\varphi\wedge \psi$ where $\varphi$ is the conjunction of literals and $\psi$ is the conjunction of linear literals without length operation. For each string variable $x$, we first compute pre-images of all linear literal $i=|x|$. Then we generate CEFAs of all regular expressions $\regex$ in all regular literals $x\in \regex$ and intersect these CEFAs and pre-images to get the final automaton $\aut_x$ for $x$. At present, there is a final CEFA for each string variable $x$. The linear arithmetic constraint $\psi$ restricts the accepting word of final CEFAs. Thus the satisfiability problem is the emptiness checking problem of final CEFAs under $\psi$. If it is empty, the string constraints are unsatisfiable. Otherwise, the string constraints are satisfiable.
% The emptiness checking problem of CEFAs under linear integer arithmetic is theoretically pspace-complete\cite{atva2020}. To solve it efficiently for a practical example, we develop heuristic ways such as under-approximation (Section \ref{subsec:under-approx}) and symbolic-aware simplification (Section \ref{subsec:simplify}).

\subsection{Pre-iamge}
\denghang{TODO}

\subsection{Emptiness Checking} \label{subsec:emptiness}
As mentioned, the emptiness checking problem of CEFAs under linear integer arithmetic $P$ is theoretically Pspace-complete. In our previous research  \cite{atva2020}, we rewrote CEFAs to an infinite system and used a model-checking tool \emph{nuXmv}\cite{nuxmv} to solve it. However, it needs to be more effective. So we put forward a new framework that brings in under-approximation and symbolic-aware simplification heuristics.\newline
As shown in Algorithm \ref{alg:emptiness}, we simplify the input CEFAs at line 1 because they may have many transitions and registers. The purpose of the simplification is to deal with duplication. After simplification, we try to find a solution by under-approximation at line 3 and line 4. An off-the-shelf SMT solver gives the result of $\varphi_{under}\wedge P$. If we find a solution, we return $false$ at line 5 because it implies that the CEFAs under linear integer arithmetic $P$ are not empty. Otherwise, we compute the Parikh images of the simplified CEFAs and check if the Parikh images are satisfiable in conjunction with $P$ at line 12. The Parikh image checking is complete so that satisfiability implies non-emptiness directly. \newline
The under-approximation program is executed many times until the $MaxBound$ is reached. $MaxBound$ is an empirical bound set to $15$ in our experiment because it brings exponential growth of searching space. The details of simplification are shown in Algorithm \ref{alg:simplify} and under-approximation are shown in Algorithm \ref{alg:underApprox}.\newline
\begin{algorithm}[h]
  \caption{ $\algfun{isEmpty}(auts, P)$}
  \label{alg:emptiness}
  \begin{algorithmic}[1]
    \Require the CEFAs $auts$ and the linear integer arithmetic $P$
    \Ensure $true$ or $false$
    \Statex
    \State $simpliAuts \leftarrow \algfun{simplify}(auts)$
    \For{$bound \gets 1, MaxBound$ }
    \State $\varphi_{under}\gets \algfun{underApprox}(simpliAuts, bound)$
    \If{$\varphi_{under}\wedge P$ is sat}
    \State \textbf{return} $false$
    \EndIf
    \EndFor
    \State $\varphi \gets \algfun{parikhImage}(simpliAuts)$
    \If{$\varphi\wedge P$ is unsat}
    \State \textbf{return} $true$
    \Else
    \State \textbf{return} $false$
    \EndIf
  \end{algorithmic}
\end{algorithm}

\subsubsection{Symbolic-Aware Simplification} \label{subsec:simplify}
The purpose of simplification is to remove duplicated transitions and registers. In the emptiness checking Algorithm \ref{alg:emptiness}, the vectors on the transitions are meritorious, while the letters are not. So we see the alphabet of the CEFA as unary (i.e., the alphabet is $\{a\}$) and see the vectors in the CEFA as letters in the NFA. The vector $\myvec{0}_n$ is seen as $\epsilon$. Precisely, given an CEFA $\aut = (Q, \Sigma, \delta, q_I, F, R, \theta)$, we obtain a symbolic form  $\aut_{sym} = (Q, \Sigma', \delta', q_I, F, R, \theta)$ where
\begin{itemize}
  \item $\Sigma' = \{a\}$,
  \item the transition set $\delta'$ is composed of transition $q\xrightarrow[\myvec{v}]{a}q'$ if there is a transition $q\xrightarrow[\myvec{v}]{b} q'$ in $\delta$ for $b\in \Sigma$ .
\end{itemize}
We directly apply algorithms of epsilon-closure, determination, and simplification in \cite{aut_hopcraft}. We see $q\xrightarrow[\myvec{0}_n]{a}q'$ as epsilon transition and compute the epsilon-closure based on it. We see $q\xrightarrow[\myvec{v}]{a}q'$ as a transition with label $(a, \myvec{v})$ and determine the automaton based on the label. Two accepting states are equivalent if they have the same accepting conditions, and two states are equivalent if they reach equivalent states for each label. We simplify the automaton by merging equivalent states.
The process above comprises line 3 and line 4 of Algorithm \ref{alg:simplify}. From line 5 to line 13, we check whether some registers are duplicated. (i.e., the values of these registers are always the same). If two registers $r_i$ and $r_j$ are duplicated at line 6, we remove one at line 7 and the corresponding update of vectors at line 9 to ensure the vectors are consistent with the new registers. After deleting the duplicated registers, we add constraint $r_i = r_j$ to store the value of the deleted register at line 11.
\begin{algorithm}[h]
  \caption{$\algfun{simplify}(auts)$}
  \label{alg:simplify}
  \begin{algorithmic}[1]
    \Require A set of CEFAs
    \Ensure The simplified CEFAs
    \Statex
    \For{$\aut \in auts$}
    \State Suppose that $\aut = (Q, \Sigma, \delta, q_I, F,
      R, \theta)$ and $R=(r_1,\cdots, r_n)$
    \State $\algfun{determinizeByVec}(\aut)$
    \State $\algfun{minimizeByVec}(\aut)$
    \ForAll{$(i, j)$ where $1\leq i \leq j \leq n$}
    \If {$\myvec{v}[i] = \myvec{v}[j]$ for every vector $\myvec{v}\in \aut$ }
    \State $R=(\cdots r_{j-1}r_{j+1}\cdots)$ \Comment remove $r_j$ from the register vector $R$
    \ForAll{$\myvec{v'}\in \aut$}
    \State $\myvec{v'}=(\cdots\myvec{v'}[j-1]\myvec{v'}[j+1]\cdots)$ \Comment remove $\myvec{v'}[j]$ from each vector $\myvec{v'}$
    \EndFor
    \State $\theta = \theta\wedge r_i = r_j$
    \EndIf
    \EndFor
    \EndFor
    \State \textbf{return} $auts$
  \end{algorithmic}
\end{algorithm}
\begin{example}
  Considering the CEFA recognizing $(a|b)\{1,100\}$ illustrated in Fig.\ref{subfig:simp_origin_cefa}. We first obtain a symbolic CEFA by using unary alphabet (Fig.\ref{subfig:simp_symbolic_cefa}) and then minimize it based on the symbolic label "$a/(1)$"(Fig.\ref{subfig:simp_final}).
  \begin{figure}[h]
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=1cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=$1\leq r \leq 100$, accepting/.style=accepting by arrow,
        ]

>>>>

As mentioned in Section \ref{sec:overview}, the main idea of our approach is to model the counting operators symbolically by registers in automata, instead of unfolding them explicitly. Additionally, we represent string lengths using registers. This reduces the satisfiability of string constraints that involve regex-counting and string-length to the satisfiability of linear integer arithmetic, which can be solved by off-the-shelf SMT solvers. We also propose techniques to reduce automata sizes and utilize under approximations to enhance performance.
\subsection{Encode Regex-Counting to CEFA} \label{subsec:regex2cefa}
One key point of our approach is the encoding of counting operator. In this section, we define the construction of CEFA from the regular expression with counting. We divide the counting operator into \emph{handled} and \emph{non-handled}. The counting operator is called \emph{non-handled} if it is the sub-regex of other complement operator (e.g., $(\regex^{\{m,n\}})^C$), closure operator(e.g., $(\regex^{\{m,n\}})^*$), or counting operator (e.g., $(\regex^{\{m,n\}})^{\{m', n'\}}$), otherwise it is called \emph{handled}. A regex is called \emph{non-handled} if it contains any non-handled counting operators, and is called \emph{handled} otherwise.
% We must syntactically rewrite bounded repetition $\regex\{m,n\}$ if it is the sub-regex of complement (e.g., $(\regex\{m,n\})^C$), closure(e.g., $(\regex\{m,n\})^*$), and bounded repetition(e.g., $(\regex\{m,n\})\{m', n'\}$), we unwind it to $\regex^m\mid\cdots\mid\regex^n$. After this syntactic rewriting, we call the resulting regex $\regex'$ \emph{non-nested}.
Given a handled regex $\regex$, operations such as intersection, union, concatenation, complement, closure, and counting may appear. To construct the CEFA of $\regex$, we first create CEFA for each sub-regex and then combine them. Similar to the construction of NFA from regex\cite{aut_hopcraft}, our construction is inductive on the size of $\regex$. The base case is the construction for a single character, an empty string, or an empty language. The inductive step is the construction for the concatenation, union, intersection of two automata of sub-regexes, counting, closure, and complement of one automaton of sub-regex. We only discuss the inductive step for counting because it is the most important and the other operations are trivial. \newline
For a handled regular expression $\regex\{m,n\}$, suppose the CEFA recognizing $\regex$ is $\aut = (Q,\Sigma,\delta,q_I,F,\emptyset,\top)$, then the CEFA recognizing $\regex\{m,n\}$ is defined as $\aut_{m,n} = (Q, \Sigma, \delta',q_ I, F, r, \theta)$ where:
\begin{itemize}
  \item $r$ is a new register,
  \item $\delta'$ is composed by transition $q\xrightarrow[(0)]{a} q'$ for each transition $q\xrightarrow[()]{a} q' \in \delta$, and transition $q_f\xrightarrow[(1)]{\epsilon} q_I$ for each accepting state $q_f\in F$,
  \item $\theta$ is the function mapping each accepting state to a same linear integer arithmetic $m\leq r\leq n$.
\end{itemize}
The construction is illustrated in Fig.\ref{fig:construct_repetition}, where a new register $r$ is added to store the counting times. The transition $q\xrightarrow[(1)]{\epsilon} q_I$ is used to update the counting times for each $q\in F$. The accepting condition $\theta'$ constrains the counting times of the accepting word to the range $[m, n]$. 
\begin{figure}[h]
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \begin{tikzpicture}[
      shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
      initial text=, every state/.style={minimum size = 0.001cm},
      accepting text=$\top$, accepting/.style=accepting by arrow,
      ]

      \node[state,initial]            (q_0)                      {};
      \node[state, accepting]         (q_1) [right=of q_0]       {};
      \node [fit=(q_0) (q_1)] {$\cdots$};
    \end{tikzpicture}
    \caption{The CEFA recognizing $\regex$}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \begin{tikzpicture}[
      shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
      initial text=, every state/.style={minimum size = 0.001cm},
      accepting text=\myemph{$\scriptstyle m\leq r\leq n$}, accepting/.style=accepting by arrow,
      ]

      \node[state,initial]            (q_0)                      {};
      \node[state, accepting]         (q_1) [right=of q_0]       {};
      \path[->] (q_1) edge[bend right]   node    [above] {\myemph{$\scriptstyle \mathbf{\epsilon:(1)}$}} (q_0);
      \node [fit=(q_0) (q_1)] {$\cdots$};
    \end{tikzpicture}
    \caption{The CEFA recognizing $\regex^{\{m,n\}}$}
  \end{subfigure}
  \caption{The construction of counting operator}
  \label{fig:construct_repetition}
\end{figure}
\begin{example}
  To construct the CEFA recognizing $(ab)\{1,100\}$. As shown in Fig.\ref{subfig:rep_aut_ab1_100}, the base cases are the constructions of character $a$ (Fig.\ref{subfig:rep_aut_a}) and $b$ (Fig.\ref{subfig:rep_aut_b}). The first inductive step is to concatenate them to one automaton (Fig.\ref{sub@subfig:rep_aut_ab}). The second inductive step is to add counting information on the automaton. To do that we add a new register $r_1$ and a transition $q_2\xrightarrow[(1)]{\epsilon} q_0$ to update the repetition times. The accepting condition in accepting state is $1\leq r_1\leq 100$ so that the counting time of the accepting word is in the range $[1,100]$. After that we get the CEFA recognizing $(ab)\{1,100\}$ shown in Fig.\ref{subfig:rep_aut_ab1_100}.
  \begin{figure}[h]
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=$\scriptstyle \top$, accepting/.style=accepting by arrow,
        accepting where=above
        ]

@@

        \path[->] (q_0) edge              node      [above]           {$\scriptstyle a:()$} (q_1);
      \end{tikzpicture}
      \caption{The CEFA recognizing $a$}
      \label{subfig:rep_aut_a}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=$\scriptstyle \top$, accepting/.style=accepting by arrow,
        accepting where=above
        ]

@@

        \path[->] (q_0) edge              node      [above]           {$\scriptstyle b:()$} (q_1);
      \end{tikzpicture}
      \caption{The CEFA recognizing $b$}
      \label{subfig:rep_aut_b}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=$\scriptstyle\top$, accepting/.style=accepting by arrow,
        accepting where=above
        ]

@@

        \path[->] (q_0) edge              node      [above]           {$\scriptstyle a:()$} (q_1)
        (q_1) edge              node      [above]           {$\scriptstyle b:()$} (q_2);
      \end{tikzpicture}
      \caption{The CEFA recognizing $ab$}
      \label{subfig:rep_aut_ab}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=$\scriptstyle\myemph{1}\leq r_1\leq\myemph{100}$, accepting/.style=accepting by arrow,
        accepting where=above
        ]

@@

        \path[->] (q_0) edge              node      [above]           {$\scriptstyle a:(\myemph{0})$} (q_1)
        (q_1) edge              node      [above]           {$\scriptstyle b:(\myemph{0})$} (q_2)
        (q_2) edge [bend left]  node      [below]           {$\scriptstyle \epsilon:(\myemph{1})$} (q_0);
      \end{tikzpicture}
      \caption{The CEFA recognizing $(ab)\{\myemph{1},\myemph{100}\}$}
      \label{subfig:rep_aut_ab1_100}
    \end{subfigure}
    \caption{The example of constructing regex with counting}
  \end{figure}
\end{example}
Above discusses the construction of counting operator in a handled regex. The counting operator of unhandled regex can not be constructed by the same way. For example, for an unhandled regex $(ab^{\{1,100\}})*$, there are infinite repetitions of regex $ab^{\{1,100\}}$. The counting times of each repetition are irrelevant, so that we can not represent them use finite registers. So we syntactically rewrite the unhandled counting operator into union operator and concatenation operator. More exactly, the sub-regex $\regex^{\{m,n\}}$ in the unhandled regex is rewritten into $\underbrace{\regex\cdot\regex\cdots\regex}_{m \text{ times}}\mid\cdots\mid \underbrace{\regex\cdot\regex\cdots\regex}_{n \text{ times}}$.
% Based on the decision procedure defined in the paper \cite{atva2020}, we propose an efficient algorithm to solve ESL formula $\varphi\wedge \psi$ where $\varphi$ is the conjunction of literals and $\psi$ is the conjunction of linear literals without length operation. For each string variable $x$, we first compute pre-images of all linear literal $i=|x|$. Then we generate CEFAs of all regular expressions $\regex$ in all regular literals $x\in \regex$ and intersect these CEFAs and pre-images to get the final automaton $\aut_x$ for $x$. At present, there is a final CEFA for each string variable $x$. The linear arithmetic constraint $\psi$ restricts the accepting word of final CEFAs. Thus the satisfiability problem is the emptiness checking problem of final CEFAs under $\psi$. If it is empty, the string constraints are unsatisfiable. Otherwise, the string constraints are satisfiable.
% The emptiness checking problem of CEFAs under linear integer arithmetic is theoretically pspace-complete\cite{atva2020}. To solve it efficiently for a practical example, we develop heuristic ways such as under-approximation (Section \ref{subsec:under-approx}) and symbolic-aware simplification (Section \ref{subsec:simplify}).
\subsection{Pre-iamge of String-length}
Given a string-length constraint, we build a CEFA with a register storing the length of accepted word. We call such a CEFA the \emph{pre-image} of it. Formally, the pre-image of length constraint is a CEFA $(\{q_I\}, \Sigma, \{q_I\xrightarrow[(1)]{\Sigma} q_I\}, q_I, \{q_I\}, r, \theta)$ where accpeting condition $\theta$ is based on the length value. For example, example \ref{eg:pre_len} shows the pre-image of length constraints $i=|x|$, in which $\theta(q_I)$ is $r = i$. This automaton contains a simple loop which increase $1$ on register $r$ for all char in $\Sigma$. So the accepting words of it is all words with length $i$. \newline
% The formal definition of pre-images of abundant string functions is listed in \cite{atva2020}, and we do not repeat it here.
\begin{example}[The pre-image of $i = |x|$]\label{eg:pre_len} 
  \begin{figure}[h]
    \centering
    \begin{tikzpicture}[
      shorten >=1pt,node distance=2cm,on grid,>={Stealth[round]},
      initial text=, every state/.style={minimum size = 0.001cm},
      accepting text=$\scriptstyle{r=i}$, accepting/.style=accepting by arrow
      ]

      \node[state,initial,accepting]            (q_0)       {$q_I$};

      \path[->] (q_0) edge [loop below] node{$\scriptstyle\Sigma:(1)$} ();
    \end{tikzpicture}
  \end{figure}
\end{example} 
\subsection{Combine Regex-counting and String-length}
In the previous section, we have discussed the construction of regex-counting operator and the pre-image of string-length constraint. In this section, we show how to combine them to deal with the string constraints with regex-counting and string-length. 
\begin{definition} [The Product of Two CEFA]\newline
  Given two CEFAs $\aut_1 = (Q_1, \Sigma, \delta_1, q_{I_1}, F_1, R_1, \theta_1)$ and $\aut_2 = (Q_2, \Sigma, \delta_2, q_{I_2}, F_2, R_2, \theta_2)$ with $R_1\cap R_2 = \emptyset$, the product of them, denoted by $\aut_1 \times \aut_2$, is defined as $(Q_1\times Q_2, \Sigma, \delta, (q_{I_1}, q_{I_2}), F_1\times F_2, R_1\cup R_2, \theta)$ where
  \begin{itemize}
    \item $\delta$ comprises the tuples $((q_1, q_2), a, (q_1', q_2'), \myvec{v_1}\cdot \myvec{v_2})$ such that $(q_1, a, q_1', \myvec{v_1})\in \delta_1$ and $(q_2, a, q_2', \myvec{v_2})\in \delta_2$.
    \item $\theta((q_1, q_2))$ is $\theta_1(q_1)\wedge \theta_2(q_2)$ for each accepting state $(q_1, q_2)\in F_1\times F_2$.
  \end{itemize}
\end{definition}
With the help of product, we can combine the CEFAs of regex-counting and string-length. For example, the CEFA of string constraints $x\in \regex \wedge i=|x|$ is the product of the CEFA of $x\in \regex$ and the pre-image of $i=|x|$.

\subsection{Emptiness Checking Problem on CEFAs} \label{subsec:emptiness}
As mentioned, the emptiness checking problem of CEFAs under linear integer arithmetic $P$ is theoretically Pspace-complete. In our previous research  \cite{atva2020}, we rewrote CEFAs to an infinite system and used a model-checking tool \emph{nuXmv}\cite{nuxmv} to solve it. However, it needs to be more effective. So we put forward a new framework that brings in under-approximation and symbolic-aware simplification heuristics.\newline
As shown in Algorithm \ref{alg:emptiness}, we simplify the input CEFAs at line 1 because they may have many transitions and registers. The purpose of the simplification is to deal with duplication. After simplification, we try to find a solution by under-approximation at line 3 and line 4. An off-the-shelf SMT solver gives the result of $\varphi_{under}\wedge P$. If we find a solution, we return $false$ at line 5 because it implies that the CEFAs under linear integer arithmetic $P$ are not empty. Otherwise, we compute the Parikh images of the simplified CEFAs and check if the Parikh images are satisfiable in conjunction with $P$ at line 12. The Parikh image checking is complete so that satisfiability implies non-emptiness directly. \newline
The under-approximation program is executed many times until the $MaxBound$ is reached. $MaxBound$ is an empirical bound set to $15$ in our experiment because it brings exponential growth of searching space. The details of simplification are shown in Algorithm \ref{alg:simplify} and under-approximation are shown in Algorithm \ref{alg:underApprox}.\newline
\begin{algorithm}[h]
  \caption{ $\algfun{isEmpty}(auts, P)$}
  \label{alg:emptiness}
  \begin{algorithmic}[1]
    \Require the CEFAs $auts$ and the linear integer arithmetic $P$
    \Ensure $true$ or $false$
    \Statex
    \State $simpliAuts \leftarrow \algfun{simplify}(auts)$
    \For{$bound \gets 1, MaxBound$ }
    \State $\varphi_{under}\gets \algfun{underApprox}(simpliAuts, bound)$
    \If{$\varphi_{under}\wedge P$ is sat}
    \State \textbf{return} $false$
    \EndIf
    \EndFor
    \State $\varphi \gets \algfun{parikhImage}(simpliAuts)$
    \If{$\varphi\wedge P$ is unsat}
    \State \textbf{return} $true$
    \Else
    \State \textbf{return} $false$
    \EndIf
  \end{algorithmic}
\end{algorithm}
\subsubsection{Symbolic-Aware Simplification} \label{subsec:simplify}
The purpose of simplification is to remove duplicated transitions and registers. In the emptiness checking Algorithm \ref{alg:emptiness}, the vectors on the transitions are meritorious, while the letters are not. So we see the alphabet of the CEFA as unary (i.e., the alphabet is $\{a\}$) and see the vectors in the CEFA as letters in the NFA. The vector $\myvec{0}_n$ is seen as $\epsilon$. Precisely, given an CEFA $\aut = (Q, \Sigma, \delta, q_I, F, R, \theta)$, we obtain a symbolic form  $\aut_{sym} = (Q, \Sigma', \delta', q_I, F, R, \theta)$ where
\begin{itemize}
  \item $\Sigma' = \{a\}$,
  \item the transition set $\delta'$ is composed of transition $q\xrightarrow[\myvec{v}]{a}q'$ if there is a transition $q\xrightarrow[\myvec{v}]{b} q'$ in $\delta$ for $b\in \Sigma$ .
\end{itemize}
We directly apply algorithms of epsilon-closure, determination, and simplification in \cite{aut_hopcraft}. We see $q\xrightarrow[\myvec{0}_n]{a}q'$ as epsilon transition and compute the epsilon-closure based on it. We see $q\xrightarrow[\myvec{v}]{a}q'$ as a transition with label $(a, \myvec{v})$ and determine the automaton based on the label. Two accepting states are equivalent if they have the same accepting conditions, and two states are equivalent if they reach equivalent states for each label. We simplify the automaton by merging equivalent states.
The process above comprises line 3 and line 4 of Algorithm \ref{alg:simplify}. From line 5 to line 13, we check whether some registers are duplicated. (i.e., the values of these registers are always the same). If two registers $r_i$ and $r_j$ are duplicated at line 6, we remove one at line 7 and the corresponding update of vectors at line 9 to ensure the vectors are consistent with the new registers. After deleting the duplicated registers, we add constraint $r_i = r_j$ to store the value of the deleted register at line 11.
\begin{algorithm}[h]
  \caption{$\algfun{simplify}(auts)$}
  \label{alg:simplify}
  \begin{algorithmic}[1]
    \Require A set of CEFAs
    \Ensure The simplified CEFAs
    \Statex
    \For{$\aut \in auts$}
    \State Suppose that $\aut = (Q, \Sigma, \delta, q_I, F,
      R, \theta)$ and $R=(r_1,\cdots, r_n)$
    \State $\algfun{determinizeByVec}(\aut)$
    \State $\algfun{minimizeByVec}(\aut)$
    \ForAll{$(i, j)$ where $1\leq i \leq j \leq n$}
    \If {$\myvec{v}[i] = \myvec{v}[j]$ for every vector $\myvec{v}\in \aut$ }
    \State $R=(\cdots r_{j-1}r_{j+1}\cdots)$ \Comment remove $r_j$ from the register vector $R$
    \ForAll{$\myvec{v'}\in \aut$}
    \State $\myvec{v'}=(\cdots\myvec{v'}[j-1]\myvec{v'}[j+1]\cdots)$ \Comment remove $\myvec{v'}[j]$ from each vector $\myvec{v'}$
    \EndFor
    \State $\theta = \theta\wedge r_i = r_j$
    \EndIf
    \EndFor
    \EndFor
    \State \textbf{return} $auts$
  \end{algorithmic}
\end{algorithm}
\begin{example}
  Considering the CEFA recognizing $(a|b)\{1,100\}$ illustrated in Fig.\ref{subfig:simp_origin_cefa}. We first obtain a symbolic CEFA by using unary alphabet (Fig.\ref{subfig:simp_symbolic_cefa}) and then minimize it based on the symbolic label "$a/(1)$"(Fig.\ref{subfig:simp_final}).
  \begin{figure}[h]
    \begin{subfigure}[b]{0.49\textwidth}
      \centering
      \begin{tikzpicture}[
        shorten >=1pt,node distance=1cm,on grid,>={Stealth[round]},
        initial text=, every state/.style={minimum size = 0.001cm},
        accepting text=$1\leq r \leq 100$, accepting/.style=accepting by arrow,
        ]
